---
title: "Urban Fledglings"
Author: "Anja Morgenthaler"
Date: Today
toc: true
abstract-title: "Research question"
abstract: "How do environmental stressors and resource availability influence reproductive success?"
format:
  html:
    embed-resources: true
    number-sections: true
    code-fold: show
    
editor: visual
editor_options:
  chunk_output_type: console
execute:
  warning: false
  message: false
---

# Urban Fledglings - Analysis Report

## Session Preparation

The preparation includes loading of the required libraries, importing the cleaned data and adding a additional column to number the nests. Then, nests with NA values are excluded for further analysis.

```{r}
#| label: Loading libraries
#| echo: false
#load libraries
library(pastecs)
  library(ggpubr); theme_set(theme_bw())
  library(GGally)
  library(performance)
  library(MuMIn)
  library(ggeffects)
  library(knitr)
  library(dplyr)
```

```{r}
#| label: Importing data
#| echo: false

birds <- read.csv("/Users/anjam/Desktop/Fledglings07/data/processed/clean.csv")
summary(birds)

birds$nests <- 1:nrow(birds)


```

```{r}
#| label: excluding NA data
#| echo: false

birds <- na.omit(birds) #clean command to delete all rows that contain NA values
birds <- birds %>%
  select(-starts_with("error"), -("nests"))
#use summary to look at data and see how many NA datapoints there are:
summary(birds)


```

## Data Exploration

First the data is being explored both numerically and graphically.

### Numerically

```{r}
#| label: tbl-explore-data
#| tbl-cap: "Descritive statistics for the variables."
#| echo: false

sum_stats <- stat.desc(birds)
kable(sum_stats, digits=1)
```

### Graphically

```{r}
#| label: fig-scatterplot-matrix
#| fig-cap: "Relationship between Insect Abundance and Environmental noise on number of offspring"
#| echo: false


ggplot(birds, aes(InsectAbundance, EnvironmentalNoise)) +
           geom_point(aes(size = Number.of.Fledglings))

```

@fig-scatterplot-matrix displays the size of offspring in relation to the environmental noise and the food abundance. There's already a trend visible, indicating a higher number of Fledglings in environments with low noise levels and high insect abundance.

## Fit Models

To analyse the data, a matching model has to be fitted. A poisson GLM is being chosen.
```{r}
#| label: Models
#| Output: false
#fit poisson models
glm.1 <- glm(Number.of.Fledglings ~1, family = "poisson", birds)
glm.2 <- glm(Number.of.Fledglings ~ InsectAbundance + EnvironmentalNoise, family = "poisson", birds)
glm.3 <- glm(Number.of.Fledglings ~ InsectAbundance * EnvironmentalNoise, family = "poisson", birds)

```

```{r}
#| label: tbl-fit-models
#| tbl-cap: "Model comparison"
#| echo: false
#Compare models
kable(compare_performance(glm.1, glm.2, glm.3))
compare_performance(glm.1, glm.2, glm.3)
```

Based on the compared performance, model glm.2 is chosen, as it shows the best fit with lowest AIC/AICc and BIC (@tbl-fit-models).

### Validate Preferred Model
The preferred model is then validated by checking the overdispersion and deviance. 
```{r}
#| label: fig-diagnostics-overdispersion
#| echo: false

check_overdispersion(glm.2)
deviance(glm.2) / df.residual(glm.2) #bad if deviance > 1
#check_model(glm.2)

```

As there's no overdispersion detected and the deviance is low, analysis will be continued with model 2.

## Interpretation

```{r}
#| label: interpret-objects
#| echo: false

#Objects we need to construct table:
#coefficients, SE, z and p
coef.table <- round(summary(glm.2)$coefficients, 3)
#approximate R2
r.2 <- round(attributes(r.squaredLR(glm.2))$adj.r.squared, 3)
#compare performance to null-model
null.chi <- anova(glm.1, glm.2, test = "Chisq") 

```

### Prediction

```{r}
#| label: fig-prediction-plot
#| echo: false


pred <- ggpredict(glm.2, terms = c("InsectAbundance", "EnvironmentalNoise"))
plot(pred)
```

#### Check prediction

```{r}
#| label: fig-checking-prediction
#| echo: false
birds$pred <- predict(glm.2, type = "response")

ggplot(birds, aes(pred, Number.of.Fledglings)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  theme_bw()
```

The prediction in @fig-checking-prediction seems to fit the actually observed values well.

```{r}
#| label: fig-predictions
#| echo: false
  pred_insect <- ggpredict(glm.2, terms = "InsectAbundance")
  pred_env <- ggpredict(glm.2, terms = "EnvironmentalNoise")
ggarrange(
  plot(pred_insect),
  plot(pred_env)
)
```

When looking at the two seperate values in @fig-predictions, there seems to be a stronger influence of Insect Abundance on the number of offspring than the environmental noise. Both show a positive correlation though.

```{r}
#| label: tbl-anova
#| echo: false
kable(anova(glm.2))
```

Also the anova of the model shows that Insect abundance has a significant influence on number of Fledglings, while the effect of the level of environmental noise is not significant!

##Store session information

```{r}
#| echo: false
sessionInfo()
```

sessionInfo()
